### 2025-06-13 conteudo/introduction/intro.tex
update research image
update intro to fit more with text and keep the style of intro of paper

The Linux kernel is a Free/Libre/Open-Source project providing value to
billions of people daily, representing 96.3\% of the top one million web
servers and 85\% of smartphones worldwide. The magnitude that the project has
reached over the years makes the addition of features, as well as handling bugs
and security vulnerabilities, potentially affect a significant number of
people, demonstrating the importance of maintaining the kernel
\citep{linuxdata}.

Maintaining the Linux kernel is a complex task, as the project contains more
than 28 million lines of code and has had over twenty thousand contributors
throughout its lifespan
\footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
Counted  28783390 lines of code and 20699 number of contributors in May 29th, 2025.
Number of contributors counted as number of unique full name listed as author of a patch.}.
Given the magnitude of the Linux
kernel project, it is unrealistic to expect that all contributors have
consistently followed the best programming practices and principles in every
contribution made to the project. Consequently, it is expected that a
significant amount of low-quality code artifacts will be found in the source
code, which could hinder or compromise the maintenance of the project and the
addition of new features.

As stated by \citet{driverdef}, device drivers are a ``piece of software whose
aim is to control and manage a particular hardware device, hence the name
device driver.'' Device drivers are a major part of the Linux kernel,
representing 66\% of the source code, and are responsible for managing
essential elements such as keyboards, mice, and GPUs \citep{marcelo}. When the
kernel opts to support new hardware, it necessitates alterations to existing
device drivers or the creation of new ones, making maintaining device drivers a
critical duty of the Linux community.

The AMD Display driver in the Linux kernel is responsible for enabling AMD GPUs
to operate correctly in desktop Linux environments, which is an essential task,
given that AMD GPUs represented 19\% of the personal GPU market in 2023
\citep{gpumarket}. When we interacted with the maintainers of the AMD Display
driver, they shared with us their challenges regarding one of the
characteristics of low-quality code artifacts: duplicated code, which hampers
the maintenance of the driver.

In searching for solutions to the AMD Display driver problem related to
duplicated code artifacts, we found tools capable of determining, given two
code artifacts, whether they are duplications of each other. However, the
literature on code duplication mainly focuses on optimizing the accuracy
of these tools for this specific task. These solutions do not address our
problem, as we cannot provide a complete codebase and have the tools return the
code duplications within it. Moreover, they do not guide us on how to mitigate
the code duplications after detection, necessitating the exploration of
alternative solutions.

Given the unsuccessful results of our search in the literature, we
turned to explore related tools not present in literature for solutions, but were also met 
with limited success.
The best solutions we found were primitive, merely
identifying every pair of code files that are duplications. The code files in
the AMD Display driver containing duplicated code typically have specific code
unique to each file. Therefore, a solution that identifies duplication in a
deeper context is necessary to find accurate duplications within the Linux
kernel context. 
The list of related tools examined in this exploratory research 
is available in Appendix \ref{app:gray}.

@@

Our studies and observations reveal that existing tools cannot fully address the code duplication 
problem described in the previous section in the formal or grey literature. To fill this gap,
we propose an approach to identify and mitigate code duplications, specifically within the
Linux kernel subsystems, focusing our study on the AMD Display driver and the Industrial 
IO subsystem. Our approach includes the development of a tool capable of detecting duplicated 
functions and establishing strategies for addressing these duplications. We tested the approach 
in the context of two kernel components, the AMD Display driver and the Industrial IO subsystem.

\begin{figure}[h]
\includegraphics[scale=0.46]{research_design}
\caption{Diagram of the research design.}
\label{fig:reDesign}
\end{figure}

This research aims to identify patterns of software engineering techniques, 
such as refactoring methods commonly used to address code duplications. 
These patterns will serve as a guide for future contributors in mitigating 
code duplication issues within this context. To investigate our approach, 
we are developing a tool to identify duplicate code. We then refactor portions 
of this code and submit patches to maintainers, providing evidence of the tool’s 
accuracy in detecting duplicate functions. Our research design follows a 
multi-method approach, structured into two phases, each employing different research methods.

In \textbf{Phase I}, we propose, develop, and validate a tool capable of
identifying code duplications at the function level within C codebases.
We expect this tool to effectively detect code duplications in the targeted subsystems 
of the Linux kernel, as preliminary investigations suggest that kernel code files often 
contain a mix of duplicated and unique functions. We will use a triangulation method to 
validate the tool, comparing its results with an established code duplication database 
referenced in the literature and conducting an empirical analysis on a randomly selected 
set of files from the AMD Display driver, which our tool flags as containing duplications.
A detailed explanation of our validation process is provided in Chapter \ref{cha:method}.

Upon completing the first phase, we anticipate having a validated tool for identifying 
code duplications within the context of the Linux kernel subsystems studied. We then 
proceed to the second phase, which focuses on mitigating these duplications. 
In \textbf{Phase II}, we aim to verify if we can mitigate the duplication found by 
the developed tool. To achieve this, we have designed an ethnographic approach that 
enables us to interact with the Linux development community and gather artifacts that 
either support or challenge our proposed mitigation strategies. Figure \ref{fig:reDesign} 
summarizes the research steps proposed in this study.

>>>>

The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, 
powering 96.3\% of the top one million web servers and 85\% of smartphones worldwide. 
Maintaining the kernel is an enormous task, involving more than 28 million lines of 
code and contributions from over twenty thousand developers
\footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
Counted  28783390 lines of code and 20699 number of contributors in May 29th, 2025.
Number of contributors counted as number of unique full name listed as author of a patch.}.
Due to the scale of the project, not all contributions adhere to best practices, resulting 
in poor-quality code artifacts that may complicate maintenance and future feature development.

One such artifact is code duplication, which is the focus of this research. The motivation for 
this work began with a practical problem rooted in device drivers, which are a major part 
of the kernel, representing 66\% of the source code\citep{marcelo}. We interacted with the 
maintainers of the AMD Display driver, an essential component for the 19\% of personal GPUs 
manufactured by AMD in 2023\citep{gpumarket}. They shared their challenges with a significant 
amount of duplicated code that hampers the driver's maintenance. In searching for a solution, 
we found that existing tools are not well-suited for identifying duplications in large-scale 
codebases like the Linux kernel, nor do they offer guidance on how to resolve the duplications 
they find. Despite searching both formal and grey literature, existing solutions fell short of 
addressing these specific challenges. The list of related tools examined in this exploratory 
research is available in Appendix \ref{app:gray}.

This research presents ArKanjo, a command-line tool designed to detect and analyze function-level 
duplications in the Linux kernel. Released under the MIT license, ArKanjo uses a two-stage 
architecture with a Preprocessor and a Query Responder. This design separates the heavy 
computational analysis from the fast querying of duplication results, making it effective for 
large codebases. We validated the tool by comparing its results with the BigCloneBench dataset
\citep{bigclonebench} and by conducting an empirical analysis on a randomly selected set of 
files from the AMD Display driver.

Beyond detection, we also investigated if the duplications found by ArKanjo could be mitigated.
This was achieved through an ethnographic approach, which included a participant observation 
study where we attempted to mitigate duplications, and a non-participant observation study of 
university students tasked with fixing duplications found by the tool. These studies showed that 
ArKanjo can effectively lower the barrier for new contributors, as evidenced by its role in 
guiding students to make their first code improvements to the kernel.

@@

To address the code duplication problem within the Linux kernel, this research employs a 
multi-method approach designed to bridge the gap between detection and practical mitigation. 
Our approach, illustrated in Figure \ref{fig:reDesign}, is structured into two interconnected phases that progress 
from tool development to an in-depth study of its application within the kernel’s development community. 
The overall objective is to produce not only a functional tool but also actionable insights into 
the patterns and practices surrounding code duplication and its resolution.

\begin{figure}[h]
\includegraphics[scale=0.9]{research_design}
\caption{Diagram of the research design.}
\label{fig:reDesign}
\end{figure}

The initial phase of our research focused on the development and validation of ArKanjo, a 
command-line tool capable of identifying code duplication at the function level.  The 
tool's accuracy was rigorously validated by evaluating its performance against the 
standard BigCloneBench dataset \citep{bigclonebench} and by conducting an empirical analysis 
of its results within the AMD Display driver.  This foundational work ensured we had a 
reliable instrument to proceed with the practical investigation of code duplication 
in our target subsystems.

With a validated tool, the research then moved from detection to practice, investigating 
the real-world challenges of mitigating the identified duplications. This was accomplished 
through an ethnographic study designed to engage with the kernel community and understand 
their perspectives on code quality and refactoring.  This study had two components: a 
participant-observation experiment where the author submitted refactoring patches to the 
AMD Display driver, and a non-participant observation of university students using ArKanjo 
to contribute fixes to both the AMD Display and Industrial I/O subsystems.

By combining tool 
development with direct community interaction, this research design allows for a comprehensive 
understanding of the technical and social factors that influence the mitigation of duplicated 
code in a large-scale open-source project.
